---
title: "The need to assistance in use of e-infrastructure within the Social sciences and Humanities"
subtitle: "at Faculty of Social Sciences at the University of Oslo."
author: "Dr. Athanasia M. Mowinckel"
affiliation: "Center for Lifespan Changes in Brain and Cognition, PSI, SV, UiO"
editor: visual
execute:
  message: false
  warning: false
  echo: true
format:
  html:
    theme: capro-quarto.scss
    toc: true
    toc-location: left
    code-fold: true
    code-summary: "Show the code"
    embed-resources: true
    fig-responsive: true
    fig-cap-location: margin
reference-location: margin
citation-location: margin
params:
  run_nettskjema: false
---

```{r 'data'}
#| include: false
source(here::here("R", "utils.R"))
ret <- get_data(315246, online = params$run_nettskjema)
dt <- ret$dt
meta <- ret$meta
```

```{r 'summaries'}
#| include: false
dt <- dt |> 
    mutate(
      institute = case_when(
      grepl("promenta", institute, ignore.case = TRUE) ~ "Promenta",
      grepl("psychology", institute, ignore.case = TRUE) ~ "PSI",
      grepl("ritmo", institute, ignore.case = TRUE) ~ "RITMO",
      grepl("psi", institute, ignore.case = TRUE) ~ "PSI",
      TRUE ~ str_trim(institute)
  ))

faculty <- dt |> 
  group_by(faculty) |> 
  tally() |> 
  mutate(pc = n/sum(n)) |> 
  arrange(pc)

inst <- dt |> 
  group_by(institute) |> 
  tally() |> 
  mutate(pc = n/sum(n)) |> 
  arrange(pc)

help <- dt |> 
  group_by(future_help_budget) |> 
  tally()
```


## Background

`r meta$elements$details[[1]]`

`r meta$elements$details[[2]]`

`r meta$elements$details[[3]]`

`r meta$elements$details[[4]]`


## Summary

On `r dateform(meta$created)` the survey "*`r meta$title`*" was launched among researchers at the University of Oslo.
As of `r dateform(as.Date(Sys.Date()))` the survey has collected `r nrow(dt)` responses from `r nrow(inst)` institutions, and among these responses, `r help |> filter(future_help_budget == "yes") |> pull(n)` said they would budget for the use of the services proposed if they were available.
Another `r help |> filter(future_help_budget == "possibly") |> pull(n)` would consider it given their needs and the skills advertised.
The need for assistance in the use of e-infrastructure is clearly signalled by the responses provided, and show as lack of support to properly utilise e-infrastructure by the researchers, either by lack of such support, or by under availability.

## Survey results

Of the `r nrow(dt)` responses to the survey, the majority of responders (`r percent(last(faculty$pc))`) were from `r last(faculty$faculty)`, and are mainly within `r last(inst$institute)` (`r percent(inst |> filter(institute %in% c("PSI", "Promenta", "Ritmo", "LCBC")) |> summarise(pc = sum(pc)) |> pull(pc))`).
Full presentation of main work place can be found in @fig-work.
The survey was disseminated by ways of the psi sympa mailing list to all scientific staff at PSI, in addition to targeted e-mails to colleagues outside of PSI that we knew might be interested in the services CAPRO aims at providing.

```{r}
#| label: fig-work
#| fig-cap: "Summary of respondents places of work, and their general field of research"
dt |> 
  mutate(
    institute = if_else(is.na(institute), "Undisclosed", institute),
    institute = fct_infreq(institute)
    ) |> 
  ggbar(institute)

```

### Past needs

A substantial part of the responders indicated they in the past needed help in utilising new technology to get their projects running (`r filter(dt, past_needs == "yes") |> nrow()`).
These report mainly asking other researchers for help, but some report asking internal center staff (RITMO, LCBC) and IT (local or central).
As seen in @fig-help, most reporting getting "some" help, which was the lowest category or possible support to receive (next to not receiving any).

```{r}
#| label: fig-help
#| fig-cap: "Help received and needed"
p1 <- dt |> 
  drop_na(past_help) |> 
  mutate(past_help = if_else(past_help == "no", "none", past_help)) |> 
  ggbar(past_help) +
  labs(title = "Amount of received help in the past")

p2 <- dt |> 
  ggplot(aes(past_satisfaction, group = past_satisfaction)) + 
  geom_bar(aes(fill = past_satisfaction), show.legend = FALSE) +
  labs( y = "Count",
        title = "Satisfaction with received help",
        x = "Rating") +
  scale_x_continuous(limits = c(0, 10.5)) +
  coord_flip()

p1 + p2

```

\

### Future needs

The survey further probed what type of needs researchers have for e-infrastructure in their projects.
These are summarised in @fig-needs, where all categories of services CAPRO can provide were selected by a minimum of `r dt |> separate_rows(future_help, sep = ";") |> group_by(future_help) |> tally() |> pull(n) |> min()` responders. 

```{r}
#| label: fig-needs
#| fig-cap: Areas researchers need help in were probed in 8 categories (in addition to an extra freetext option). Each bar is its own grouping, showing the percent of responders that indicated they needed help with this specific category.

dt |> 
  separate_rows(future_help, sep = ";") |> 
  group_by(future_help) |> 
  tally() |> 
  mutate(
    pc = n/nrow(dt),
    future_help = fct_infreq(future_help)
  ) |>
  ggplot(aes(future_help, pc)) +
  geom_bar(stat = "identity", aes(fill = pc), 
           alpha = .7, 
           color = "grey82", 
           show.legend = FALSE) +
  geom_label(aes(label = scales::percent(pc)), 
             nudge_x = .1, 
             alpha = .5) +
  labs(title = "") +
  coord_polar(clip = "off") +
  scale_y_continuous(labels = scales::percent) +
  theme(axis.title.x = element_blank())
```

## Conclusion

The need and interest for assistance in setting up and utilising e-infrastrucure exists among a lot of researchers at PSI in particular, but likely also in other parts of UiO and external institutions. 
CAPRO could provide valuable assistance to more efficiently and more transparently work with data through custom and standardised applications, as well as through workshops. 

