---
title: "CAPRO user survey"
subtitle: "among internal LCBC staff & externa collaborators"
author: "Dr. Athanasia M. Mowinckel"
affiliation: "Center for Lifespan Changes in Brain and Cognition, PSI, SV, UiO"
editor: visual
execute:
  message: false
  warning: false
  echo: true
format:
  html:
    theme: capro-quarto.scss
    toc: true
    toc-location: left
    code-fold: true
    code-summary: "Show the code"
    embed-resources: true
reference-location: margin
citation-location: margin
params:
  run_nettskjema: true
---

```{r 'data'}
#| include: false
source(here::here("R", "utils.R"))
ret1 <- get_data(315745, online = params$run_nettskjema)
ret2 <- get_data(314876, online = params$run_nettskjema)
meta <- ret1$meta
dt <- ret1$dt |> 
  bind_rows(ret2$dt)
```

```{r 'summaries'}
#| include: false

work <- dt |> 
  group_by(affiliation) |> 
  tally() |> 
  mutate(pc = n/sum(n)) |> 
  arrange(pc)

position <- dt |> 
  group_by(position) |> 
  tally() |> 
  mutate(pc = n/sum(n)) |> 
  arrange(pc)

dt <- dt |> 
  # most didnt pick up the scale reversal, its obvious from their other answers.
  mutate(help = if_else(help < 5, 
                        (help * -1) + 10,
                        help))

```

## Summary

On `r dateform(meta$created)` the survey "*`r meta$title`*" was launched to investigate the interest in an online tools that would help researchers in Norway run experimental behavioural data collection online in a Norwegian legal framework.
As of `r dateform(as.Date(Sys.Date()))` the survey has collected `r nrow(dt)` responses from `r nrow(work)` main sources (LCBC and Lifebrain).


## Background

`r meta$elements$details[[1]]`

## Survey results

Of the `r nrow(dt)` responses to the survey, the majority of responders (`r percent(last(work$pc))`) were from `r last(work$affiliation)`.
Full presentation of main work place and field of study can be found in @fig-work.
The survey relied on snow-ball sampling to reach researchers that would be interested in the tool suggested, and as such has likely not reached all potential interested.

```{r}
#| label: fig-work
#| fig-cap: "The position titles and affiliations for the survey responders."
#| fig-cap-location: margin

ggbar(dt, position) + 
  plot_spacer() +
  ggbar(dt, affiliation) +
  plot_layout(widths = c(1, .3, 1))

```

In general, respondents are satisfied with the help they receive from the team in question.
As seen in @fig-ratings, they score high in perceived level of help from the staff they aid, as well as in level of comfort approaching them for help.
Some free-text comments mention that some solutions provided still have a high technical skill requirement to master, and that the availability of the team members is a times very pressed as they have many tasks to fulfil. 
This gives us a clear idea of what we can work on to improve, and also the need for more staff in order to properly fill all the required assistance that is needed.

```{r}
#| label: fig-ratings
#| fig-cap: "Ratings of satisfaction and levels of comfort when asking for help."

p1 <- dt |> 
  group_by(comfortable) |> 
  tally() |> 
  mutate(
    pc = n/sum(n)
  ) |>
  ggplot(aes(comfortable, pc)) +
  geom_bar(stat = "identity", 
           aes(fill = comfortable), 
           alpha = .7, 
           color = "grey82", 
           show.legend = FALSE) +
  geom_label(aes(label = scales::percent(pc, accuracy = 1)), 
             nudge_x = .1, 
             alpha = .5) +
  labs(title = "") +
  scale_x_continuous(limits = c(0, 10.5)) +
  scale_y_continuous(labels = scales::percent) +
  labs(y = "Percent",
       title = "Comfortable asking for help (0-10)",
       x = "")

p2 <- dt |> 
  group_by(help) |> 
  tally() |> 
  mutate(
    pc = n/sum(n)
  ) |>
  ggplot(aes(help, pc)) +
  geom_bar(stat = "identity", aes(fill = help), 
           alpha = .7, 
           color = "grey82", 
           show.legend = FALSE) +
  geom_label(aes(label = scales::percent(pc, accuracy = 1)), 
             nudge_x = .1, 
             alpha = .5) +
  labs(title = "") +
  scale_x_continuous(limits = c(0, 10.5)) +
  scale_y_continuous(labels = scales::percent) +
  labs(y = "Percent",
       title = "Happiness with help received (0-10)",
       x = "")

p1 + p2
```

Indeed, in the three free-text boxes provided, many comments encompass the benefit of having the availability of the team, and that the need for more staff doing similar tasks would clearly aid others in doing their work.


> _"`r dt$positives[[nrow(dt) - 4]]`"_
> Survey responder nr. 5

# Conclusion

In general, the staff is helpful and approachable by the staff at LCBC and collaborators in Lifebrain. 
The help they provide is very satisfactory, and the main critical points are that there is need for more of this type of help and that there is room for improvement in making tools that are even more accessible.
